Pre-trained encoders for dialogue: The case of laughter

### 2019-12-09 reports/ACL2020 Paper/acl2020-paper.tex
table -> Table; figure -> Figure

Dialogue, especially spoken dialogue, is radically different from the kind of data that neural language models are pre-trained on.
In the example from the Switchboard corpus, shown in table~\ref{table:example}, it is evident that the syntactic structure of dialogue is very different from  written text, even if one imagines a script for a movie or a play.
Not only is the internal structure of contributions different---with features such as disfluencies, repair, incomplete sentences, and various vocal sounds---but the sequential structure of the discourse is also very different.
In dialogue, speakers take turns and switch perspectives.
More generally, the way that utterances contribute to and cohere with the discourse is different from the relationship between sentences in a piece of written text.

@@

\paragraph{Transfer learning for NLP}
Transfer learning techniques allow a model trained on one task---often unsupervised---to be applied to another. 
Since annotating natural language data is expensive, there is a lot of interest in transfer learning for natural language processing. 
Word vectors \citep[e.g.,][]{mikolovDistributedRepresentationsWords2013,penningtonGloveGlobalVectors2014} are a ubiquitous example of transfer learning in NLP.
We note, however, that pre-trained word vectors are not always useful when applied to dialogue domains \citet{cerisaraEffectsUsingWord2vec2017}.

@@

To test the effectiveness of BERT for DAR, we employ a simple neural architecture with two components: an encoder that vectorizes utterances, and a sequence model that predicts dialogue act tags from the vectorized utterances (figure \ref{fig:model-architecture}).
Since we are primarily interested in comparing different utterance encoders, we use a basic RNN as the sequence model in every configuration. 
The RNN takes the encoded utterance as input at each time step,
and its hidden state is passed to a simple linear classification layer over dialogue act tags.
Conceptually, the encoded utterance represents the context-agnostic semantic content of the utterance, and the hidden state of the RNN represents the full discourse context.

@@

\section{Experiments}
\subsection{Experiment 1: Impact of laughter} \label{sec:experiment1}   % Vlad
In the first experiment we investigated whether laughter, as an example of a dialogue-specific signal, is a helpful feature for DAR.
Therefore, we train another version of each model: one containing laughs (\texttt{L}) and one with laughs left out (\texttt{NL}), and compare their performances in DAR task.
Table~\ref{table:laughter-total-acc} compares the results from applying the models with two different utterance encoders (\texttt{BERT}, \texttt{CNN}).
First of all, we can see that \texttt{BERT} utterance encoder outperforms \texttt{CNN} in all the cases.
From observing the effect of laughters, we can see differences in performance depending on the dialogue act, regardless how often laughter occurs in the current or adjacent utterances (see figure~\ref{fig:by-da} in Appendix~\ref{sec:suppl}).
The strongest evidence for laughter as a helpful feature was found in SWDA for the \texttt{BERT} utterance encoder, where macro-F1 score increases by 7.89 percentage points.

@@

Confusion matrices (figure \ref{fig:swda-cm}) provide some food for thought. Most of the misclassifications fall into the majority classes, such as \emph{sd} (Statement-non-opinion), in the left edge of the matrix. However there are some important exceptions, such as \emph{rhetorical questions}, that are misclassified as other forms of questions due to their surface question-like form.
Importantly, laughter helpes to classify rhetorical questions correctly, this is because in a conversation it can be used as a device to cancel seriousness \citep{ginzburg2015understanding}, in this case, of a question.
Therefore questions, like the one we show in the example in table~\ref{table:example-qh}, are easier to disambiguate with the help of laughter.
\begin{table}
      \small
  \centering
  \begin{tabularx}{\linewidth}{llL}
    \toprule
    Speaker & DA & Utterance \\ \midrule
        \multicolumn{3}{c}{\emph{(Talking about hobbies)}}\\
    B   & sd	&  Um, as far as spare time, they talked about, \\
    B	& \% & I don't, + I think, \\
    B	& qh & who has any spare time \texttt{<laughter>}? \\
    A	& x & \texttt{<laughter>}.\\
             \bottomrule
  \end{tabularx}
  \caption{Example from the SWDA corpus (sw3735). B's contribution \emph{qh} (Rhetorical question) is misinterpreted as \emph{qw} (Wh-question) by the BERT model without laughs in training data. }
  \label{table:example-qh}
\end{table}

@@

The pre-trained model is more accurate than the randomly initialized model by several percentage points on both DA corpora,
suggesting that BERT's extensive pre-training does provide some useful information for recoginizing different dialogue acts (table \ref{table:exp2}).
This performance boost is much more pronounced in the macro-averaged F1 score,
which is explained by the fact that at the tag level, pre-training has a larger impact on less frequent tags.

@@

Comparing the performance of the fine-tuned encoder to that of the frozen model, we note that pre-training alone is not enough.
The frozen model is heavily biased towards the most frequent tags, which explains its especially poor macro-average score (table \ref{table:exp2}).
In SWDA, for example, the model with a frozen encoder predicts one of the two most common tags (Statement-non-opinion or Acknowledge) 86\% of the time, whereas those two
tags account for only 51\% of the ground truth tags.
The fine-tuned model is much less biased; it predicts the two most common tags only 59\% of the time.

@@

The results are mixed.
In-domain pre-training offers a modest boost in overall accuracy,
but the macro-average scores are almost uniformly worse (table \ref{table:exp3}).
%Indeed, when BERT is frozen during fine-tuning, the model that received no further pre-training performs better by more than 3 percentage points (table \ref{tab:exp3-recall}).
To fully assess the potential impact of further in-domain pre-training, a larger dialogue corpus is required.

>>>>

Dialogue, especially spoken dialogue, is radically different from the kind of data that neural language models are pre-trained on.
In the example from the Switchboard corpus, shown in Table~\ref{table:example}, it is evident that the syntactic structure of dialogue is very different from  written text, even if one imagines a script for a movie or a play.
Not only is the internal structure of contributions different---with features such as disfluencies, repair, incomplete sentences, and various vocal sounds---but the sequential structure of the discourse is also very different.
In dialogue, speakers take turns and switch perspectives.
More generally, the way that utterances contribute to and cohere with the discourse is different from the relationship between sentences in a piece of written text.

@@

\subsection{Transfer learning for NLP}
Transfer learning techniques allow a model trained on one task---often unsupervised---to be applied to another. 
Since annotating natural language data is expensive, there is a lot of interest in transfer learning for natural language processing. 
Word vectors \citep[e.g.,][]{mikolovDistributedRepresentationsWords2013,penningtonGloveGlobalVectors2014} are a ubiquitous example of transfer learning in NLP.
We note, however, that pre-trained word vectors are not always useful when applied to dialogue domains \citet{cerisaraEffectsUsingWord2vec2017}.

@@

To test the effectiveness of BERT for DAR, we employ a simple neural architecture with two components: an encoder that vectorizes utterances, and a sequence model that predicts dialogue act tags from the vectorized utterances (Figure~\ref{fig:model-architecture}).
Since we are primarily interested in comparing different utterance encoders, we use a basic RNN as the sequence model in every configuration. 
The RNN takes the encoded utterance as input at each time step,
and its hidden state is passed to a simple linear classification layer over dialogue act tags.
Conceptually, the encoded utterance represents the context-agnostic semantic content of the utterance, and the hidden state of the RNN represents the full discourse context.

@@

\section{Experiments}
\subsection{Experiment 1: Impact of laughter} \label{sec:experiment1}   % Vlad
In the first experiment we investigated whether laughter, as an example of a dialogue-specific signal, is a helpful feature for DAR.
Therefore, we train another version of each model: one containing laughs (\texttt{L}) and one with laughs left out (\texttt{NL}), and compare their performances in DAR task.
Table~\ref{table:laughter-total-acc} compares the results from applying the models with two different utterance encoders (\texttt{BERT}, \texttt{CNN}).
First of all, we can see that \texttt{BERT} utterance encoder outperforms \texttt{CNN} in all the cases.
From observing the effect of laughters, we can see differences in performance depending on the dialogue act, regardless how often laughter occurs in the current or adjacent utterances (see Figure~\ref{fig:by-da} in Appendix~\ref{sec:suppl}).
The strongest evidence for laughter as a helpful feature was found in SWDA for the \texttt{BERT} utterance encoder, where macro-F1 score increases by 7.89 percentage points.

@@

Confusion matrices (Figure~\ref{fig:swda-cm}) provide some food for thought. Most of the misclassifications fall into the majority classes, such as \emph{sd} (Statement-non-opinion), in the left edge of the matrix. However there are some important exceptions, such as \emph{rhetorical questions}, that are misclassified as other forms of questions due to their surface question-like form.
Importantly, laughter helpes to classify rhetorical questions correctly, this is because in a conversation it can be used as a device to cancel seriousness \citep{ginzburg2015understanding}, in this case, of a question.
Therefore questions, like the one we show in the example in Table~\ref{table:example-qh}, are easier to disambiguate with the help of laughter.
\begin{table}
      \small
  \centering
  \begin{tabularx}{\linewidth}{llL}
    \toprule
    Speaker & DA & Utterance \\ \midrule
        \multicolumn{3}{c}{\emph{(Talking about hobbies)}}\\
    B   & sd	&  Um, as far as spare time, they talked about, \\
    B	& \% & I don't, + I think, \\
    B	& qh & who has any spare time \texttt{<laughter>}? \\
    A	& x & \texttt{<laughter>}.\\
             \bottomrule
  \end{tabularx}
  \caption{Example from the SWDA corpus (sw3735). B's contribution \emph{qh} (Rhetorical question) is misinterpreted as \emph{qw} (Wh-question) by the BERT model without laughs in training data. }
  \label{table:example-qh}
\end{table}

@@

The pre-trained model is more accurate than the randomly initialized model by several percentage points on both DA corpora,
suggesting that BERT's extensive pre-training does provide some useful information for recoginizing different dialogue acts (Table~\ref{table:exp2}).
This performance boost is much more pronounced in the macro-averaged F1 score,
which is explained by the fact that at the tag level, pre-training has a larger impact on less frequent tags.

@@

Comparing the performance of the fine-tuned encoder to that of the frozen model, we note that pre-training alone is not enough.
The frozen model is heavily biased towards the most frequent tags, which explains its especially poor macro-average score (Table~\ref{table:exp2}).
In SWDA, for example, the model with a frozen encoder predicts one of the two most common tags (Statement-non-opinion or Acknowledge) 86\% of the time, whereas those two
tags account for only 51\% of the ground truth tags.
The fine-tuned model is much less biased; it predicts the two most common tags only 59\% of the time.

@@

The results are mixed.
In-domain pre-training offers a modest boost in overall accuracy,
but the macro-average scores are almost uniformly worse (Table~\ref{table:exp3}).
%Indeed, when BERT is frozen during fine-tuning, the model that received no further pre-training performs better by more than 3 percentage points (Table~\ref{tab:exp3-recall}).
To fully assess the potential impact of further in-domain pre-training, a larger dialogue corpus is required.
